---
title: "Ignoring Missing Values"
author: "Robert Horton"
date: "November 7, 2014"
output: ioslides_presentation
---

## Missing Values

By default, many R functions will not give you an answer if there are missing values in the input:
```{r mean_with_NA}
measurements <- c(1,3,2,NA,10)

mean(measurements)
```
Since it doesn't know what the fourth value should have been, the default result is to not give you an answer.

## Dropping Missing Values

We can specify that we want the function to ignore the missing values, and just operate on the values it has,  by setting the `na.rm` parameter to TRUE.

```{r mean_with_na_rm}
measurements
mean(measurements, na.rm=TRUE)
```

## Applying a Function to a Matrix
```{r apply_a_function}
A <- matrix(1:12, nrow=4)
A
apply(A, 1, mean)
```
The 1 tells apply to use the function on the first dimension of the matrix (the rows). To find the means of the columns, you could use
`apply(A, 2, mean)`.

## Applying a Function to a Matrix
Note that R has built-in functions for `rowMeans` and `colMeans`:
```{r rowMeans}
apply(A, 1, mean)
rowMeans(A)
```
These functions are faster than using `apply`, which can make a big difference with large data sets.

## Applying a Function to a Matrix
However, `apply` is more general; you can use it with any vectorized function, including those you write yourself.
```{r appl_custom_function}
maxMinusMin <- function(v) max(v) - min(v)
apply(A, 1, maxMinusMin)
```

## Missing Values in Matrixes
What if there are NA values in our data matrix?
```{r matrix_B}
B <- A
B[1,3] <- NA
B
apply(B, 1, mean)
```

How do we tell `mean` to ignore NAs?

## Passing parameters to applied functions
The elipses construct lets you refer to any parameters not otherwise specified:
```{r ellipses}
maxMinusMin <- function(v, ...) max(v, ...) - min(v, ...)
apply(A, 1, maxMinusMin, na.rm=TRUE)
```
Here the three dots (`...`) represent any additional parameters passed to the function; this makes it easy to pass these parameters on to subfunctions. In this case, we will call both `min` and `max` with `na.rm` set to `TRUE`, since that argument was passed to `maxMinusMin`.

Use the help function (e.g., `?mean`) to see the parameters it takes.
