---
title: "The Regression Equation"
author: "Bob Horton"
date: "April 19, 2015"
output: html_document
runtime: shiny
---

This is a multivariate example of the regression equation.

## Linear regression in 3D

Here x1 and x2 are independent variables, and will be used to identify points on a 2D plane; the dependent variable will be plotted on the z-axis.


```{r, echo=FALSE}
inputPanel(
  selectInput("N", label = "Number of points:",
              choices = c(10, 25, 50, 100, 200), selected = 25),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



