---
title: 'Lecture 2a: Probability'
author: "Robert Horton"
date: "December 28, 2014"
output: ioslides_presentation
---

## Shuffling Cards
```{r car_suites, echo=FALSE, eval=FALSE}
library(Unicode)
x <- as.u_char(9824:9827)
suites <- u_char_inspect(as.u_char(9824:9827))$Char
```
```{r shuffling_cards}
deck <- paste0(c(2:10, "J", "Q", "K", "A"), rep(c("♠", "♡", "♢", "♣"), each=13))
set.seed(123)
rbind(sample(deck, 5), sample(deck, 5), sample(deck, 5), sample(deck, 5))  # wrong
set.seed(123)
deck <- sample(deck)
cbind(deck[1:4], deck[5:8], deck[9:12], deck[13:16], deck[17:20])
```


## Conditional Probability
[Bayes Theorem](http://en.wikipedia.org/wiki/Bayes%27_theorem)

$$
P(A|B) = \frac{P(B|A) P(A)}{P(B)} 
$$

Special case for two sets A and B:

$$
P(A|B)=\frac{P(B|A)P(A)}{P(B|A)P(A)+P(B|\neg A)P(\neg A)}
$$

## Diagnostic tests: sensitivity and specificity
    T  = positive test
    T' = negative test
    D  = disease
    D' = no disease

    sensitivity: P(T|D)
    specificity: P(T'|D')
    prevalence of disease: P(D)
    post-test probability: P(D|T)
        a.k.a "positive predictive value"
        
##
```{r, echo=FALSE}
# http://en.wikipedia.org/wiki/Positive_and_negative_predictive_values#Relationship
```

<STYLE type="text/css">    
table, th, td {
    border: 1px solid black;
    font-size: 1.1em;
    text-align: center;
    padding-top: 5px;
    padding-bottom: 5px;
}
tr.alt td {
    color: #000000;
    background-color: #EAF2D3;
}
</STYLE>

|    | D                | D'                | Prevalence = D/(D+D')   |             |
|----|------------------|-------------------|-------------------------|-------------|
| T  | TP               | FP                | precision = PPV = TP/T  | FDR = FP/T  |
| T' | FN               | TN                | FOR = FN/T'             | NPV = TN/T' |
|    | sensitivity=TP/D | FPR=FP/D'         | accuracy=(TP+TN)/(D+D') |             |
|    | FNR=FN/D         | specificity=TN/D' |                         |             |

(From [Wikipedia](http://en.wikipedia.org/wiki/Positive_and_negative_predictive_values#Relationship))

$$ sensitivity = \frac{number~of~true~positives}{number~with~disease} $$
    
$$ specificity = \frac{number~of~true~negatives}{number~without~disease} $$
    
$$ prevalence = \frac{number~with~disease}{total~population} $$


## Exercise
Given a screening test with 

* sensitivity of 98%, 
* specificity of 95%, and 
* disease prevalence of 0.1%, 

what is the probability that a person who tests positive actually has the disease?

## Apply Bayes Theorem
    P(T|D) = 0.98
    P(T'|D') = 0.95
    P(D) = 0.001
    
    P(D|T) = P(T|D)*P(D)/P(T)
    
    P(T) = P(T|D)*P(D) + P(T|D')P(D')
    P(T|D')P(D') = (1 - P(T'|D')) * (1 - P(D))
    
    P(D|T) = P(T|D)*P(D)/( P(T|D)*P(D) + (1 - P(T'|D')) * (1 - P(D)) )


## R Code
```{r positive_predictive_value}
post_test_probability <- function(sensitivity, specificity, prevalence)
	sensitivity*prevalence/(sensitivity*prevalence + (1-specificity)*(1-prevalence))

post_test_probability(sensitivity=0.98, specificity=0.95, prevalence=0.001)
```


## Applications of medical tests
* [screening](http://en.wikipedia.org/wiki/Type_I_and_type_II_errors#Medical_screening): detect or predict disease in at-risk population.
* [diagnosis](http://en.wikipedia.org/wiki/Medical_test#Diagnostic): confirm or determine disease in a person suspected of having it.
* [monitoring](http://en.wikipedia.org/wiki/Monitoring_(medicine)): measurement of conditions over time.

## Permutations: order matters

Permutation with replacement
$$n^k$$

Permutation without replacement
$${k!\,(n-k)!}$$

## Combinations: order does not matter
Combinations (order doesn't matter)
$$\tbinom nk = \tfrac{n!}{k!\,(n-k)!}$$

```{r permutations}
choose(n=3, k=1)
choose(n=3, k=2)
choose(n=10, k=3)
choose(52,5)
```

## Assignment
* Write a program to estimate the probability of getting two pair in 5-card stud poker. Compare your answer to the probability given in the Wikipedia article listing [poker hands](http://en.wikipedia.org/wiki/List_of_poker_hands).